---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# ThermalSampleR

`ThermalSampleR` is an R package and R Shiny GUI application designed for assessing sample size requirements for researchers performing critical thermal limits (CTL) studies (e.g. calculating CTmin or CTmax metrics). Much research has been performed in recent years to improve the methodology used during CTL studies, however, we are not aware of any research into the sample size requirements for these studies. Our package allows users to perform sample size assessments for both single-species studies and multi-species comparisons, which will be discussed in detail below and illustrated with full use-cases. 

`ThermalSampleR` is a companion package to the pending publication (currently under review): 

Owen, C.A., Sutton, G.F., Martin, G.D., van Steenderen, C.J.M., and Coetzee, J.A. Sample size planning for critical thermal limits studies. 


## 1. Installation

**Via GitHub:**       
`devtools::install_github("CJMvS/ThermalSampleR")`

Once the package has been installed, you need to call the package into your current R session: 

```{r}
library(ThermalSampleR)
```

**R Shiny Application:**      

**Via GitHub:**      
`shiny::runUrl("https://github.com/CJMvS/ThermalSampleR_Shiny/archive/main.tar.gz")`   

**Shiny Apps platform:**    
https://clarkevansteenderen.shinyapps.io/ThermalSampleR_Shiny/    

## 2. Loading your raw data 

A full worked example will be outlined in the following sections. The first step is to load in your raw critical thermal limits raw data. Input files must be saved in .csv format, with two columns: one column containing unique species names (indicated by the `col` column below) and another column containing the response variable, with each row representing a single individual that has been tested (e.g. Critical Thermal Limit temperature data) (indicated by the `response` column below). For example, you can inspect the built-in example data (`coried`) in `ThermalSampleR` to see how your data should be structured:  

```{r load_example_data}
coreid <- ThermalSampleR::coreid_data
head(coreid)
```

## 3. Sample size assessments - Single taxon 

The simplest application of `ThermalSampleR` is to evaluate whether a study has used a sufficient sample size to estimate the critical thermal limits for a single taxon. Below, we perform these calculations to estimate sample sizes required to accurately estimate the CTmin of adults of the a twig-wilting insect *Catorhintha schaffneri* (denoted by `Catorintha schaffneri_APM` in our dataframe). 

We use a bootstrap resampling procedure to estimate the width of the 95% confidence interval of our CTmin estimate across a range of sample sizes, which defaults to starting at n = 3 individuals tested, and which can be extrapolated to sample sizes greater than the sample size of your existing data by specifying a value to `n_max`. 

```{r boot_one_sim, cache = TRUE}
bt_one = boot_one(
  # Which dataframe does the data come from? 
  data = coreid_data, 
  # Provide the column name containing the taxon ID
  groups_col = col, 
  # Provide the name of the taxon to be tested
  groups_which = "Catorhintha schaffneri_APM", 
  # Provide the name with the response variable (e.g CTmin data)
  response = response, 
  # Maximum sample sample to extrapolate to 
  n_max = 49, 
  # How many bootstrap resamples should be drawn? 
  iter = 299)
head(bt_one)
```

The resulting variable should then be passed to the `plot_one_group` function to visualise the simulation results. A number of optional parameters can be passed to the function to alter the aesthetics of the graphs. 

```{r boot_one_plot, fig.width=14, fig.height=6}
plot_one_group(
  # Variable containing the output from running `boot_one` function
  x = bt_one, 
  # Minimum sample size to plot
  n_min = 3, 
  # Actual size of your existing dataset 
  n_max = 15, 
  # Colour for your experimental data
  colour_exp = "forestgreen", 
  # Colour for the extrapolated predictions 
  colour_extrap = "orange", 
  # Position of the legened
  legend.position = "right", 
  # Change the degree of shading on the graph 
  alpha_val = 0.25)
```

Inspecting panel (a), we visualise the precision of our CTmin estimate for *Catorhintha schaffneri*, whereby precision is measured as the width of a 95% confidence interval. For example, in the context of CTLs, a CI width of 1 indicates that practitioners can be 95% confident that their CTL estimate is within 1 degree Celsius of the true CTmin value. The smaller the CI width, the greater the precision of the CTL estimate.  

In this example, the precision of our CTmin estimate was high and is not predicted to improve substantially by increasing sample size, once approximately n = 20 individuals are tested, as the 95% confidence interval reaches a plateau at n = 20. The plateau is in the extrapolation section of the graph indicating that more individuals would need to be tested for the 95% confidence interval to become approximately stable. However, at the existing sample size of n = 15, the researchers could be relatively confident that the CTmin estimate they have obtained is precise to within approximately 1.2 - 1.5 degrees Celsius. Researchers will need to decide for themselves what an acceptable degree of precision is for their own datasets. 

Inspecting panel (b), we visualise the sampling distribution (i.e. the range of plausible CTmin values) for the taxa under study. This assessment can produce bias results at small sample sizes because the population parameter (e.g. the taxons' CTmin) is unknown, and therefore must be estimated from the experimental data. Panel B gives an indication of parameter estimation accuracy by plotting the proportion of bootstrap resamples, across each sample size, for which the 95% CI included the estimated population parameter. An accurate parameter estimate should produce CI's that, on ~ 95% of occasions, contain the estimated population parameter. 

In this example, the accuracy of our CTmin estimate was high once n > 10 individuals were tested. The proportion of 95% CI's containing the estimated population parameter  approximated the expectation that 95% of the CI's for a given sample size should ideally contain the population parameter once n = 10 were tested. As noted above, because our population parameter for *Catorhintha schaffneri* was estimated from n = 15 individuals tested, our assessment of parameter accuracy may be bias, and thus, should be interpreted with caution. 

***Take-home message***: As long as the researchers were content with obtaining a CTmin estimate for *Catorhintha schaffneri* with a precision of approximately 1.2 - 1.5 degrees Celsius, the experiment could be concluded at n = 15 individuals tested. Adding additional samples above n = 15 would likely improve the precision of the CTmin estimate, however, the gain in precision must be considered in light of the logistics, costs and ethics of testing additional specimens. 




